{\bf Practical Concurrency Testing}

### 2018-10-12 thesis/defense/snakefight.tex
write intro and concl

\newcommand\hilight[2]{\color{#1}#2\color{black}}
\definecolor{darkorange}{RGB}{192,128,0}
\definecolor{darkblue}{RGB}{0,97,192}
\definecolor{olivegreen}{RGB}{0,128,0}

@@

\begin{frame}{Challenges} %\& Existing Coping Techniques}
	Unpredictable state space explosion
	\begin{itemize}
		\item Equivalence reductions: DPOR {\em [Flanagan '05]}, MCR {\em [Huang '15]}
		\item Heuristic search ordering: ICB {\em [Musuvathi '08]}, PCT {\em [Burckhardt '10]}
		\item There will always be a {\bf bigger test case}!
	\end{itemize}
	\pause
	\linegap

	Too much manual intervention for non-expert use
	\begin{itemize}
		\item Manual kernel annotations: Landslide {\em [Blum '12]} {\tiny that's me!}
			% TODO: smth about choosing pps?
		\item State space size estimation {\em [Simsa '12]}
			% TODO change this bullet pt to be more explicitly about 
		\item Widespread adoption requires {\bf fully automated instrumentation}!
	\end{itemize}
	\pause
	\linegap

	Limited sources of nondeterminism
	\begin{itemize}
		\item Timer-driven schedule nondeterminism (all prior work)
		\item Weak memory nondeterminism {\em [Zhang '15]}
		\item Event-driven applications {\em [Jensen '15]}
		\item {\bf Hardware transactional memory}?
	\end{itemize}
\end{frame}

@@

\definecolor{gray}{RGB}{128,128,128}

\begin{frame}{Thesis Statement}
	The Iterative Deepening algorithm makes MC appropriate for both educational and real-world programs.
	\vspace{4.14em}

	Ad-hoc synchronization heuristics make MC accessible to student programmers.
	\vspace{4.14em}

	A new transactional concurrency model keeps MC applicable to the latest real-world techniques.
	\vspace{4.14em}
\end{frame}
\begin{frame}{Outline}
	\hilight{gray}{The Iterative Deepening algorithm makes MC appropriate for both educational and real-world programs.} \\
	\begin{itemize}
		\item {\bf Quicksand and Iterative Deepening}
		\begin{itemize}
			\item Algorithm overview \& data-race preemption points
			\item Verification \& experimental results
		\end{itemize}
	\end{itemize}
	\hilight{gray}{Ad-hoc synchronization heuristics make MC accessible to student programmers.} \\
	\begin{itemize}
		\item {\bf Educational use}
		\begin{itemize}
			\item Pebbles \& 15-410 user study
			\item Pintos \& upcoming plans
		\end{itemize}
	\end{itemize}

	\hilight{gray}{A new transactional concurrency model keeps MC applicable to the latest real-world techniques.} \\
	\begin{itemize}
		\item {\bf Transactional memory}
		\begin{itemize}
			\item Example \& challenges
			\item Equivalence \& failure injection
		\end{itemize}
	\end{itemize}
\end{frame}

@@

\begin{frame}{Results - Bugs found}
	\begin{center}
		\includegraphics[width=0.8\textwidth]{../../oopsla/bugs-talk.pdf}
	\end{center}
\end{frame}
\begin{frame}{Results - Full verifications}
	\begin{center}
		\includegraphics[width=0.8\textwidth]{../../oopsla/verifs-talk.pdf}
	\end{center}
\end{frame}

@@

	Otherwise? Quicksand's heuristic search ordering finds bugs faster anyway
	\pause
	\linegap

@@

\begin{frame}{Results}
	Participation
	\begin{itemize}
		\item CMU: 103/177 students (7 semesters)
		\item PSU: 38/136 students (1 semester)
		\item U. Chicago: 4/21 students (1 semester)
	\end{itemize}
	\linegap

	Bug-finding
	\begin{itemize}
		\item CMU: 115 deterministic, 122 concurrency bugs in 75/103 P2s
		\item U. Chicago: 3 deterministic, 4 concurrency bugs in 3/4 kernels
		% \item Case study tej
	\end{itemize}
\end{frame}

\begin{frame}{Effect on Grades}
	Compared Landslide users' vs. non-users' submitted P2 grades...
	\begin{itemize}
		\item ``Does Landslide help students submit better projects?''
	\end{itemize}
	\linegap

@@

\begin{frame}{Effect on Grades}
	\begin{center}
		\includegraphics[width=\textwidth]{../photo-of-ze-studence.pdf}
	\end{center}
\end{frame}

@@

\begin{frame}{Survey}
	14(ish) questions, investigate both education \& happiness
	\begin{itemize}
		\item {\em Education:} Did Landslide help students learn more from P2?
		\item {\em Happiness:} Did Landslide help students save time/effort?
	\end{itemize}
	\linegap

	Self-report any false positives
	\begin{itemize}
		\item Landslide crashes aside, 3 FP bug reports, all simple bugs (now fixed)
			% nonexistent MAGIC_BREAK bc of different address space in shell
			% simultaneous landslides corrupt each other
			% invalid heap access when new_pages "too close"
		\item No FPs related to yield-loop/etc heuristics
	\end{itemize}
\end{frame}

@@

\begin{frame}{Summary}
	Landslide helps students find \& fix bugs during projects
	\begin{itemize}
		\item 73\% of CMU users found $\ge 1$ bug
		\item 56\% of CMU users fixed $\ge 1$ bug
		\item 34\% of CMU users fixed all found bugs
	\end{itemize}
	\linegap

@@

	Survey results showed students enlightenedly happy
	\linegap

@@

\begin{frame}{Example}
	\begin{center}
		\begin{tabular}{l|l}
			{\bf Thread 1} & {\bf Thread 2} \\
			\hline
			{\tt \hilight{darkorange}{\_xbegin}() == SUCCESS} \\
			{\tt count++;} \\
				& {\tt \hilight{darkorange}{\_xbegin}() == SUCCESS} \\
				& {\tt count++;} \\
				& {\tt \hilight{darkblue}{\_xend}();} \\
			{\tt \hilight{darkblue}{\_xend}()}...? \\
			{\em <conflict detected>} \\
			\\
			{\tt \hilight{darkorange}{\_xbegin}() == FAIL} \\
			{\tt \hilight{darkorange}{mutex\_lock}(...);} \\
			{\tt count++;} \\
			{\tt \hilight{darkblue}{mutex\_unlock}(...);} \\
		\end{tabular}
	\end{center}
\end{frame}

@@

	Equivalence theorem
	\begin{itemize}
		\item Safe to automatically skip all interleavings within transactions %$\qed$
			\begin{itemize}
				\item Exponential state space reduction
				\item Traditional reduction (DPOR) applied on top
			\end{itemize}
		\item More in bonus slides...
	\end{itemize}
\end{frame}

@@

\begin{frame}{Example bug}
	\begin{center}
		\begin{tabular}{l|l}
			{\bf Thread 1} & {\bf Thread 2} \\
			\hline
			{\tt \hilight{darkorange}{\_xbegin}() == FAIL} \\
			{\tt \hilight{darkorange}{mutex\_lock}(...);} \\
			{\tt count++;} \\
			{\tt \hilight{darkblue}{mutex\_unlock}(...);} \\
				& {\tt \hilight{darkorange}{\_xbegin}() == SUCCESS} \\
				& {\tt count++;} \\
				& {\tt \hilight{darkblue}{\_xend}();} \\
			\\
			\\
		\end{tabular}
	\end{center}
\end{frame}

\begin{frame}{Example bug}
	\begin{center}
		\begin{tabular}{l|l}
			{\bf Thread 1} & {\bf Thread 2} \\
			\hline
			{\tt \hilight{darkorange}{\_xbegin}() == FAIL} \\
			{\tt \hilight{darkorange}{mutex\_lock}(...);} \\
			{\tt int tmp = count;} \\
			{\em \hilight{red}{<data-race PP>}} \\
				& {\tt \hilight{darkorange}{\_xbegin}() == SUCCESS} \\
				& {\tt count++;} \\
				& {\tt \hilight{darkblue}{\_xend}();} \\
			{\tt count = tmp + 1;} \\
			{\tt \hilight{darkblue}{mutex\_unlock}(...);} \\
		\end{tabular}
	\end{center}
\end{frame}

@@

\begin{frame}{Example bug - Solution {\em [Dehesa-Azuara '16]}}
	\begin{center}
		\begin{tabular}{l|l}
			{\bf Thread 1} & {\bf Thread 2} \\
			\hline
			{\tt \hilight{darkorange}{\_xbegin}() == FAIL} \\
			{\tt \hilight{darkorange}{mutex\_lock}(...);} \\
			{\tt int tmp = count;} \\
			{\em \hilight{red}{<data-race PP>}} \\
				& {\tt \hilight{darkorange}{\_xbegin}() == SUCCESS} \\
				& {\tt \hilight{red}{if (mutex\_held(...))}} \\
				& {\tt \hilight{red}{~~~~\_xabort();}} \\
			{\tt count = tmp + 1;} \\
			{\tt \hilight{darkblue}{mutex\_unlock}(...);} \\
				& {\tt \hilight{darkorange}{mutex\_lock}(...);} \\
				& {\tt count++;} \\
				& {\tt \hilight{darkblue}{mutex\_unlock}(...);} \\
		\end{tabular}
	\end{center}
\end{frame}

@@

\begin{frame}{Evaluation - Verification}
	$K,N$ = $K$ threads, each $N$ iterations of test logic
	\linegap

	Microbenchmarks
	\begin{itemize}
			% numbers taken from retry sets column
		\item {\tt mutex\_held()} pattern: $K,N$ = 2,4 (79min); 4,1 (6min)
		\item Using {\tt xadd} instead: $K,N$ = 2,4 (19min); 3,2 (39min), 4,1 (1min)
		\item More, see thesis...
	\end{itemize}
	\pause
	\linegap

@@

\begin{frame}{Evaluation - Verification}
	Abstraction reduction {\em [Simsa '13]}
	\begin{itemize}
		\item Verify locking primitives \& client code separately against common API
	\end{itemize}
	\linegap

	HTM locks
	\begin{itemize}
			% numbers are now using STM semantics instead
			% bc of retry loop around the xbegin
			% can't really mention this in the talk
		\item Spinlock, $K,N$ = 2,2 (2min), 3,1 (28sec)
			\begin{itemize}
				\item Depends on ad-hoc yield-loop heuristics; slow
			\end{itemize}
		\item Spin replaced w/P2 mutex: $K,N$ = 2,4 (7hrs), 4,1 (11min)
	\end{itemize}
	\pause
	\linegap

@@

\begin{frame}{Conclusion}
\end{frame}

@@

\begin{frame}{Happens-Before Example}
	\begin{center}
		\begin{tabular}{ll}
			{\bf Thread 1}  & {\bf Thread 2} \\
			\texttt{\hilight{red}{x++;}}    & \\
			\texttt{mutex\_lock(m);}	& \\
			\texttt{// unrelated}	   & \\
			\texttt{mutex\_unlock(m);}      & \\
				& \texttt{mutex\_lock(m);} \\
				& \texttt{// unrelated} \\
				& \texttt{mutex\_unlock(m);} \\
				& \texttt{\hilight{red}{x++;}} \\
		\end{tabular}
		\linegap

		No race under Pure HB; {\em true potential race} under Limited HB.
	\end{center}
\end{frame}
\begin{frame}{Happens-Before Example}
	\begin{center}
		\begin{tabular}{ll}
			{\bf Thread 1}  & {\bf Thread 2} \\
			\texttt{\hilight{red}{x++;}}    & \\
			\texttt{mutex\_lock(m);}	& \\
			\texttt{\hilight{darkblue}{y = true;}}      & \\
			\texttt{mutex\_unlock(m);}      & \\
				& \texttt{mutex\_lock(m);} \\
				& \texttt{\hilight{darkblue}{bool tmp = y;}} \\
				& \texttt{mutex\_unlock(m);} \\
				& \texttt{\hilight{darkblue}{if (tmp)}~\hilight{red}{x++;}} \\
		\end{tabular}
		\linegap

@@

\begin{frame}{Quicksand results - Bugs found by CPU time}
	\begin{center}
		\includegraphics[width=0.8\textwidth]{../../oopsla/bugs-talk.pdf}
	\end{center}
\end{frame}
\begin{frame}{Quicksand results - Bugs found by wall-clock time}
	\begin{center}
		\includegraphics[width=0.8\textwidth]{../../oopsla/bugs-wallclock-talk.pdf}
	\end{center}
\end{frame}
\begin{frame}{Quicksand results - Tests fully verified (CPU time)}
	\begin{center}
		\includegraphics[width=0.8\textwidth]{../../oopsla/verifs-talk.pdf}
	\end{center}
\end{frame}
\begin{frame}{Quicksand results - MC + DR vs single-pass DR}
	\begin{center}
		\includegraphics[width=0.8\textwidth]{../../oopsla/dr-false-negatives-poster.pdf}
	\end{center}
\end{frame}
\begin{frame}{Quicksand results - Representative of real-world programs?}
	Bugs by required number of preemptions to expose:
	\begin{center}
		\footnotesize
	\begin{tabular}{r|c|c}
		{\bf Bound} & {\bf SSS (sync PPs only)} & {\bf SSS (preempt everywhere)} \\
		\hline
		0       & 2     & 1     \\
		1       & 82    & 86    \\
		2       & 16    & 32    \\
		3       & 2     & 3     \\
		4+      & 0     & 0     \\
		\hline
		Total   & 102   & 122   \\
	\end{tabular}
	\end{center}
	\linegap

@@

\definecolor{lavender}{RGB}{102,65,208} % (?? i don't know what happened here anymore but it looks good)
\definecolor{seafoam}{RGB}{56,158,68} % * 2/3
\definecolor{salmon}{RGB}{208,89,74} % * 7/8
\definecolor{goldish}{RGB}{128,96,32}
\newcommand\tii{\ensuremath{\hilight{lavender}{\mathbf{Ti}}}\xspace}
\newcommand\tjj{\ensuremath{\hilight{seafoam}{\mathbf{Tj}}}\xspace}
\newcommand\tkk{\ensuremath{\hilight{salmon}{\mathbf{Tk}}}\xspace}

@@

>>>>

\newcommand\hilight[2]{\color{#1}#2\color{black}}
\definecolor{darkorange}{RGB}{192,128,0}
\definecolor{darkblue}{RGB}{0,97,192}
\definecolor{olivegreen}{RGB}{0,128,0}
\definecolor{lavender}{RGB}{102,65,208} % (?? i don't know what happened here anymore but it looks good)
\definecolor{seafoam}{RGB}{56,158,68} % * 2/3
\definecolor{salmon}{RGB}{208,89,74} % * 7/8
\definecolor{goldish}{RGB}{128,96,32}
\definecolor{gray}{RGB}{128,128,128}

@@

\begin{frame}{Challenges} %\& Existing Coping Techniques}
	Unpredictable state space explosion
	\begin{itemize}
		\item Equivalence reductions: DPOR {\em [Flanagan '05]}, MCR {\em [Huang '15]}
		\item Heuristic search ordering: ICB {\em [Musuvathi '08]}, PCT {\em [Burckhardt '10]}
		\item There will always be a {\bf bigger test case}...
	\end{itemize}
	\pause
	\linegap

	Too much manual intervention %for non-expert use
	\begin{itemize}
		\item Manual kernel annotations: Landslide {\em [Blum '12]} {\tiny that's me!}
			% TODO: smth about choosing pps?
		\item State space size estimation {\em [Simsa '12]}
			% TODO change this bullet pt to be more explicitly about 
		\item Widespread adoption requires {\bf fully automated instrumentation}...
	\end{itemize}
	\pause
	\linegap

	Sources of nondeterminism
	\begin{itemize}
		\item Timer-driven schedule nondeterminism (all prior work)
		\item Weak memory nondeterminism {\em [Zhang '15]}
		\item Event-driven applications {\em [Jensen '15]}
		\item {\bf Hardware transactional memory}...
	\end{itemize}
\end{frame}

@@

\begin{frame}{Thesis Statement}
	% data race pps (as well as txn equivalence)
	\hilight{gray}{Combining both theoretically-founded automatic reduction techniques
	% iterative deepening
	% (say smth like: "landslide also has a wealth of other user-tunable heuristics
	% for expert users, which you can read about in the document; for this talk i will focus just on the CPU budget thing")
	% XXX: need to prepare for the question -- "what evidence do you have that users actually want this cpu budget stuff",
	% ",or, that it helps vs just running something unlimited vs a normal icb test"
	% (well, for the 2nd half of the Q, i can talk about how ID outperforms ICB on verficiations)
	and user-informed heuristic ones, stateless model checking...}
	\begin{itemize}
		\item ...can sufficiently mitigate exponential explosion...
		% education
		\item ...to be a practical testing technique for inexperienced users...
		% htm
		% TODO: this last part you just added for this talk; does the thesis version of the statmeent need to include too?
		\item ...and real-world programs alike.
	\end{itemize}
\end{frame}

\begin{frame}{Thesis Statement}
	The Iterative Deepening algorithm %combines theory and heuristics to
	practically debugs and verifies many exponential state spaces.
	\vspace{4.54em}

	Landslide's automatic instrumentation makes MC accessible to students.
	\vspace{4.54em}

	New reduction techniques for transactional memory keep MC up-to-date
	with the latest real-world concurrency techniques.
	\vspace{4.54em}
\end{frame}
\begin{frame}{Outline}
	\hilight{gray}{The Iterative Deepening algorithm %combines theory and heuristics to
	practically debugs and verifies many exponential state spaces.} \\
	%\hilight{gray}{The Iterative Deepening algorithm makes MC appropriate for both educational and real-world programs.} \\
	\begin{itemize}
		\item {\bf Quicksand and Iterative Deepening}
		\begin{itemize}
			\item Data-race preemption points \& algorithm overview
			\item Verification \& experimental results
		\end{itemize}
	\end{itemize}
	\hilight{gray}{Landslide's automatic instrumentation makes MC accessible to students.} \\
	%\hilight{gray}{Ad-hoc synchronization heuristics make MC accessible to student programmers.} \\
	\begin{itemize}
		\item {\bf Educational use}
		\begin{itemize}
			\item Pebbles \& Pintos OS projects
			\item Bug-finding, grades, \& surveys
		\end{itemize}
	\end{itemize}

	\hilight{gray}{New reduction techniques for transactional memory keep MC up-to-date
	with the latest real-world concurrency techniques.} \\
	%\hilight{gray}{A new transactional concurrency model keeps MC applicable to the latest real-world techniques.} \\
	\begin{itemize}
		\item {\bf Transactional memory}
		\begin{itemize}
			\item Concurrency model \& failure injection
			\item Bug-finding \& verification
		\end{itemize}
	\end{itemize}
\end{frame}

@@

\begin{frame}{Results - Bugs found (CPU time)}
	\begin{center}
		\includegraphics[width=0.8\textwidth]{../../oopsla/bugs-talk.pdf}
	\end{center}
\end{frame}
\begin{frame}{Results - Bugs found (wall-clock time)}
	\begin{center}
		\includegraphics[width=0.8\textwidth]{../../oopsla/bugs-wallclock-talk.pdf}
	\end{center}
\end{frame}
\begin{frame}{Results - Full verifications}
	\begin{center}
		\includegraphics[width=0.8\textwidth]{../../oopsla/verifs-talk.pdf}
	\end{center}
\end{frame}

@@

	Otherwise? Quicksand's heuristic search ordering finds bugs faster anyway
	%\pause
	%\linegap

@@

\begin{frame}{Results}
	Participation
	\begin{itemize}
		\item CMU: 103/177 students (7 semesters)
		\item U. Chicago: 4/21 students (1 semester)
		\item PSU: 38/136 students (1 semester)
	\end{itemize}
	\linegap

	Bug-finding
	\begin{itemize}
		\item CMU: 115 deterministic, 122 concurrency bugs in 75/103 P2s
		\item U. Chicago: 3 deterministic, 4 concurrency bugs in 3/4 kernels
			% TODO: what to do abt this bullet pts
			%\pause
		\item PSU: (no log collection due to IRB issues)
		% \item Case study tej
	\end{itemize}
\end{frame}

\begin{frame}{Effect on Grades (CMU)}
	Compared Landslide users' vs. non-users' submitted P2 grades...
	\begin{itemize}
		\item ``Does Landslide help students submit better projects?''
	\end{itemize}
	\linegap

@@

\begin{frame}{Effect on Grades (CMU)}
	\begin{center}
		\includegraphics[width=\textwidth]{../photo-of-ze-studence.pdf}
	\end{center}
\end{frame}

@@

\begin{frame}{Survey}
	14(ish) questions, investigate both education \& happiness
	\begin{itemize}
		\item {\bf Education:} Did Landslide help students learn more from P2?
		\item {\bf Happiness:} Did Landslide help students save time/effort?
	\end{itemize}
	\pause
	\linegap

	Self-report any {\bf false positives}
	\begin{itemize}
		\item Landslide crashes aside, 3 FP bug reports, all simple bugs (now fixed)
			% nonexistent MAGIC_BREAK bc of different address space in shell
			% simultaneous landslides corrupt each other
			% invalid heap access when new_pages "too close"
		\item No FPs related to yield-loop/etc heuristics
	\end{itemize}
	\linegap

	Self-report opinions {\bf whether it was worthwhile}
	\begin{itemize}
		\item Mostly ``worked as advertised'' praise
		\item One PSU student found a bug from inspecting a data-race by hand,
			after Landslide timed out testing with no bug report
	\end{itemize}
\end{frame}

@@

\begin{frame}{Summary}
	Landslide helps students find \& fix bugs during projects
	% TODO: should this be moved ot the earlier bug finding slide
	\begin{itemize}
		\item 73\% of CMU users found $\ge 1$ bug
		\item 56\% of CMU users fixed $\ge 1$ bug
		\item 34\% of CMU users fixed all found bugs
	\end{itemize}
	\linegap

@@

	Survey results showed students ``enlightenedly happy'' % dave eckhardt scare quotes if i ever saw some
	\linegap

@@

\begin{frame}{Example}
	\begin{center}
		\begin{tabular}{l|l}
			{\bf Thread 1} & {\bf Thread 2} \\
			\hline
			{\tt \hilight{darkorange}{\_xbegin}() == SUCCESS} \\
			{\tt count++;} \\
				& {\tt \hilight{darkorange}{\_xbegin}() == SUCCESS} \\
				& {\tt count++;} \\
				& {\tt \hilight{darkblue}{\_xend}();} \\
			{\tt \hilight{darkblue}{\_xend}()}...? \\
			{\em <conflict detected>} \\
			\\
			{\tt \hilight{darkorange}{\_xbegin}() == ABORTED} \\
			{\tt \hilight{darkorange}{mutex\_lock}(...);} \\
			{\tt count++;} \\
			{\tt \hilight{darkblue}{mutex\_unlock}(...);} \\
		\end{tabular}
	\end{center}
\end{frame}

@@

	Equivalence theorem
	\begin{itemize}
		\item Safe to automatically skip all interleavings within transactions %$\qed$
			\begin{itemize}
				\item Exponential state space reduction
				\item Traditional reduction (DPOR) applied on top
			\end{itemize}
			% TODO prob just comment this out
		\item More in bonus slides...
	\end{itemize}
\end{frame}

@@

\begin{frame}{Example bug}
	\begin{center}
		\begin{tabular}{l|l}
			{\bf Thread 1} & {\bf Thread 2} \\
			\hline
			{\tt \hilight{darkorange}{\_xbegin}() == ABORTED} \\
			{\tt \hilight{darkorange}{mutex\_lock}(...);} \\
			{\tt count++;} \\
			{\tt \hilight{darkblue}{mutex\_unlock}(...);} \\
				& {\tt \hilight{darkorange}{\_xbegin}() == SUCCESS} \\
				& {\tt count++;} \\
				& {\tt \hilight{darkblue}{\_xend}();} \\
			\\
			\\
		\end{tabular}
	\end{center}
\end{frame}

\begin{frame}{Example bug}
	\begin{center}
		\begin{tabular}{l|l}
			{\bf Thread 1} & {\bf Thread 2} \\
			\hline
			{\tt \hilight{darkorange}{\_xbegin}() == ABORTED} \\
			{\tt \hilight{darkorange}{mutex\_lock}(...);} \\
			{\tt int tmp = count;} \\
			{\em \hilight{red}{<data-race PP>}} \\
				& {\tt \hilight{darkorange}{\_xbegin}() == SUCCESS} \\
				& {\tt count++;} \\
				& {\tt \hilight{darkblue}{\_xend}();} \\
			{\tt count = tmp + 1;} \\
			{\tt \hilight{darkblue}{mutex\_unlock}(...);} \\
		\end{tabular}
	\end{center}
\end{frame}

@@

\begin{frame}{Example bug - Solution {\em [Dehesa-Azuara '16]}}
	\begin{center}
		\begin{tabular}{l|l}
			{\bf Thread 1} & {\bf Thread 2} \\
			\hline
			{\tt \hilight{darkorange}{\_xbegin}() == ABORTED} \\
			{\tt \hilight{darkorange}{mutex\_lock}(...);} \\
			{\tt int tmp = count;} \\
			{\em \hilight{red}{<data-race PP>}} \\
				& {\tt \hilight{darkorange}{\_xbegin}() == SUCCESS} \\
				& {\tt \hilight{red}{if (mutex\_held(...))}} \\
				& {\tt \hilight{red}{~~~~\_xabort();}} \\
			{\tt count = tmp + 1;} \\
			{\tt \hilight{darkblue}{mutex\_unlock}(...);} \\
				& {\tt \hilight{darkorange}{mutex\_lock}(...);} \\
				& {\tt count++;} \\
				& {\tt \hilight{darkblue}{mutex\_unlock}(...);} \\
		\end{tabular}
	\end{center}
\end{frame}

@@

\begin{frame}{Evaluation - Verification}
	$K,N$ = $K$ threads, each $N$ iterations of test logic; 10 hour timeout
	\linegap

	Microbenchmarks
	\begin{itemize}
			% numbers taken from retry sets column
		\item {\tt mutex\_held()} pattern: $K,N$ = 2,4 (79min); 4,1 (6min)
		\item Using {\tt xadd} instead: $K,N$ = 2,4 (19min); 3,2 (39min), 4,1 (1min)
		\item Is that ``enough''? Generalize with manual inspection?
		%\item More, see thesis...
	\end{itemize}
	\pause
	\linegap

@@

\begin{frame}{Evaluation - Verification}
	Abstraction reduction {\em [Simsa '13]}
	\begin{itemize}
		\item Verify locking primitives \& client code separately against common API
			% say: again relying on human to write the right test case
		\item {\em Add} verification costs instead of {\em multiplying}
	\end{itemize}
	\linegap

	HTM locks
	\begin{itemize}
			% numbers are now using STM semantics instead
			% bc of retry loop around the xbegin
			% can't really mention this in the talk
		\item Spinlock: $K,N$ = 2,2 (2min), 3,1 (28sec)
			\begin{itemize}
				\item Depends on ad-hoc yield-loop heuristics; slow
			\end{itemize}
		\item Spin replaced w/P2 mutex: $K,N$ = 2,4 (7hrs), 4,1 (11min)
	\end{itemize}
	\pause
	\linegap

@@

\begin{frame}{Conclusion}
	% data race pps (as well as txn equivalence)
	Combining %both
	{\bf theoretically-founded automatic reduction techniques}
	\vspace{4.82em}

	and {\bf user-informed heuristic ones},
	\vspace{4.82em}

	stateless model checking...
	\begin{itemize}
		\item ...can sufficiently mitigate exponential explosion...
		\item ...to be a practical testing technique for inexperienced users...
		\item ...and real-world programs alike.
	\end{itemize}
\end{frame}
\begin{frame}{Conclusion}
	Combining %both
	{\bf theoretically-founded automatic reduction techniques}
	\begin{itemize}
		\item Verification with data-race preemption points
		\item Yield-loop blocking
		\item Equivalent emulation of HTM semantics
	\end{itemize}
	and {\bf user-informed heuristic ones},
	\begin{itemize}
		\item Quicksand's CPU budget argument
		\item Exposing data-race reports to students
		\item Abstraction reduction for HTM primitives
	\end{itemize}
	\hilight{gray}{stateless model checking...}
	\begin{itemize}
		\item \hilight{gray}{...can sufficiently mitigate exponential explosion...}
		\item \hilight{gray}{...to be a practical testing technique for inexperienced users...}
		\item \hilight{gray}{...and real-world programs alike.}
	\end{itemize}
\end{frame}
\begin{frame}{Closing remarks}
	``Verifying C programs is doomed, why can't PL design handle this?''
	\begin{itemize}
		\item seL4 {\em [Klein '09]}: secure kernel verified end-to-end in Haskell
		\item Rust {\em [Matsakis '14]}: concurrency-aware type system, no data races
	\end{itemize}
	\pause
	\linegap

	But... we still have to try.
	\begin{itemize}
		\item Computers can't handle $O(N^K)$
			\begin{itemize}
				\item ``ok, $N^K$, that's, wow, 700 years...''
			\end{itemize}
		\item Humans can help?
			\begin{itemize}
				\item ``can we get same results with smaller $N$, $K$?''
			\end{itemize}
	\end{itemize}
	\pause
	\linegap

	One day live in peace with our silicon children?
\end{frame}

@@

\begin{frame}{Happens-Before Example}
	\begin{center}
		\begin{tabular}{ll}
			{\bf Thread 1}  & {\bf Thread 2} \\
			\texttt{\hilight{red}{x++;}}    & \\
			\texttt{mutex\_lock(m);}	& \\
			\texttt{\hilight{lavender}{// unrelated}}	   & \\
			\texttt{mutex\_unlock(m);}      & \\
				& \texttt{mutex\_lock(m);} \\
				& \texttt{\hilight{lavender}{// unrelated}} \\
				& \texttt{mutex\_unlock(m);} \\
				& \texttt{\hilight{red}{x++;}} \\
		\end{tabular}
		\linegap

		No race under Pure HB; {\em true potential race} under Limited HB.
	\end{center}
\end{frame}
\begin{frame}{Happens-Before Example}
	\begin{center}
		\begin{tabular}{ll}
			{\bf Thread 1}  & {\bf Thread 2} \\
			\texttt{\hilight{red}{x++;}}    & \\
			\texttt{mutex\_lock(m);}	& \\
			\texttt{\hilight{lavender}{y = true;}}      & \\
			\texttt{mutex\_unlock(m);}      & \\
				& \texttt{mutex\_lock(m);} \\
				& \texttt{\hilight{lavender}{bool tmp = y;}} \\
				& \texttt{mutex\_unlock(m);} \\
				& \texttt{\hilight{lavender}{if (tmp)}~\hilight{red}{x++;}} \\
		\end{tabular}
		\linegap

@@

%\begin{frame}{Quicksand results - Bugs found by CPU time}
%	\begin{center}
%		\includegraphics[width=0.8\textwidth]{../../oopsla/bugs-talk.pdf}
%	\end{center}
%\end{frame}
%\begin{frame}{Quicksand results - Bugs found by wall-clock time}
%	\begin{center}
%		\includegraphics[width=0.8\textwidth]{../../oopsla/bugs-wallclock-talk.pdf}
%	\end{center}
%\end{frame}
%\begin{frame}{Quicksand results - Tests fully verified (CPU time)}
%	\begin{center}
%		\includegraphics[width=0.8\textwidth]{../../oopsla/verifs-talk.pdf}
%	\end{center}
%\end{frame}
\begin{frame}{Quicksand results - MC + DR vs single-pass DR}
	\begin{center}
		\includegraphics[width=0.8\textwidth]{../../oopsla/dr-false-negatives-poster.pdf}
	\end{center}
\end{frame}
\begin{frame}{Quicksand results - Representative of real-world programs?}
	Bugs by required number of preemptions to expose:
	\begin{center}
		\footnotesize
	\begin{tabular}{r|c|c}
		{\bf Bound} & {\bf SSS (sync PPs only)} & {\bf SSS (preempt everywhere)} \\
		\hline
		0       & 2     & 1     \\
		1       & 82    & 86    \\
		2       & 16    & 32    \\
		3       & 2     & 3     \\
		4+      & 0     & 0     \\
		\hline
		Total   & 102   & 122   \\
	\end{tabular}
	\end{center}
	\linegap

@@

\newcommand\tii{\ensuremath{\hilight{lavender}{\mathbf{Ti}}}\xspace}
\newcommand\tjj{\ensuremath{\hilight{seafoam}{\mathbf{Tj}}}\xspace}
\newcommand\tkk{\ensuremath{\hilight{salmon}{\mathbf{Tk}}}\xspace}

@@

\begin{frame}{How good is $K,N$?}
	\begin{center}
		\small
		\begin{tabular}{c||c|c|c||c}
			{\bf test} & {\bf \#txn} & {\bf \#sync} & {\bf \#race} & {\bf max events verified} \\
			\hline
			{\tt htm2}	      & 1     & 2     & 4     & 42    \\ % 2,3
			{\tt counter}	   & 1     & 0     & 1     & 16    \\ % 2,4
			{\tt swap}	      & 1     & 4     & 8     & 52    \\ % 2,2
			{\tt fig63}	     & 1     & 0     & $K$-1 & 18    \\ % 3,2 k=3 -> 18; 2,4 -> 16
			{\tt avl\_insert}       & 1     & 2     & 7     & 20    \\ % 2,1 nb. only 5 drpps under -A -S but w/e
			% nb less than 9 on stm
			{\tt avl\_fixed}	& 1     & 2     & 9     & 24    \\ % 2,1 nb. again, only 5 under -A -S
			% nb stm takes it down from 13 to 10 races
			{\tt map\_basic}	& 1     & 4     & 13    & 36    \\ % 2,1
			% mysteriously had this at 6 at one point but ive no idea where that came from
			{\tt map\_basicer}      & 1     & 2     & 5     & 32    \\ % 2,2

			% TODO: check these again, will 2,4 even complete on mutex
			{\tt lock}({\tt spinlock})      & 1     & 2     & 4     & 28    \\ % 2,2
			{\tt lock}({\tt spin\_fixed})   & 1     & 2     & 4     & 28    \\ % 2,2 or 4,1
			{\tt lock}({\tt mutex})	 & 1     & 2     & 2     & 30    \\ % 2,3 gives 30; 2,4 would give 40
			% not putting lock_fast because it like, doesn't have a "state space" per se
		\end{tabular}
	\end{center}
	\linegap

	\begin{itemize}
		\item {\em [Abdulla '14]} reports 19 threads (??)
			\begin{itemize}
				\item {\sf filesystem(19)}: classic DPOR: 4096; (their) optimal DPOR: 64
			\end{itemize}
			% not really a quote, but their x axis from figure 9 graph, whose y axis goes up to 10k
		\item {\em [Zhang '15]} verifs up to 10 ``concurrent events'' (weak mem r/w?)
	\end{itemize}
\end{frame}
